### 2023-09-25 manuscript/manuscript.Rmd
results and disucssion updates.

---
title: "Imposed form in the Early Acheulean? Evidence from Gona, Afar, Ethiopia"
author:
- Dietrich Stout^[Department of Anthropology, Emory University, Atlanta, GA, USA;  dwstout@emory.edu]
- Cheng Liu^[Department of Anthropology, Emory University, Atlanta, GA, USA; raylc1996@outlook.com]
- Antoine Muller^[Institute of Archaeology, Mount Scopus, The Hebrew University of Jerusalem, Jerusalem, Israel; antoine.muller@mail.huji.ac.il]
- Michael J. Rogers^[Department of Anthropology, Southern Connecticut State University, New Haven, CT, USA; rogersm1@southernct.edu]
- Sileshi Semaw^[Centro Nacional de Investigación sobre la Evolución Humana (CENIEH), Burgos, Spain; sileshi.semaw@cenieh.es]
abstract: 
|
  ```{=latex}
  TBD.

  \textbf{Keywords:} Gona; TBD; TBD; TBD; TBD; TBD; TBD
  ```
date: "`r Sys.Date()`"
fontsize: 11pt
output: bookdown::pdf_document2
header-includes:
   - \usepackage{textcomp}
   - \usepackage{utopia}
   - \usepackage{lineno}
   - \usepackage[utf8]{inputenc}
   - \linenumbers
   - \usepackage{float}
   - \floatplacement{figure}{H}
   - \usepackage{caption}
   - \captionsetup[figure]{font=scriptsize}
   - \captionsetup[table]{font=scriptsize}
linestretch: 1.5
link-citations: yes
linkcolor: blue
bibliography: bibliography.bib
csl: apa.csl
editor_options: 
  markdown: 
    wrap: sentence
---

# Introduction

The imposition of intended form on artifacts has long been viewed as a watershed in human cognitive and cultural evolution and is most commonly associated with the emergence of "Large Cutting Tools" (LCTs) in the Early Acheulean [@isaac1976; @kuhn2020; @holloway1969].
However, this interpretation of Acheulean LCTs as intentionally designed artifacts remains controversial.
Alternative proposals range from the possibility that LCTs were unintended by-products of flake production [@moore2016; @noble1996] to the suggestion that their form was "at least partly under genetic control" [@corbey2016].
Even accepting that LCT form was to some extent intended, there is substantial disagreement over the specificity of design.
Some analyses have indicated that shape variation in Acheulean handaxes is largely a result of resharpening [@iovita2011; @mcpherron2000] whereas others find form to be unrelated to reduction intensity and more likely to reflect normative expectations of what handaxes should look like [@garcía-medrano2019; @shipton2015; @shipton2020].
Such debates about shape of Acheulean LCTs may appear narrowly technical but have broad relevance for evolutionary questions including the origins of human culture [@corbey2016; @tennie2017; @shipton2015], language [@stout2012], teaching [@gärdenfors2017], brain structure [@hecht2015], and cognition [@stout2015; @wynn2016].
To examine these questions, we studied the complete collection of Early Acheulean flaked pieces from 5 sites at Gona Project Area and compared them with Oldowan cores from 2 published sites at Gona.
By comparing shape variation to measures of flaking intensity and patterning, we sought to identify technological patterns that might reveal intent.

## Identifying design

There is a broad consensus that refined handaxes and cleavers from the later Acheulean resulted from procedurally elaborate, skill intensive, and socially learned production strategies [@caruana2020; @garcía-medrano2019; @moore2020; @sharon2009; @shiptonEvolutionSocialTransmission2019; @stout2014] although debate over the presence of explicit, culturally transmitted shape preferences continues [@iovita2011; @moore2020; @shipton2020; @wynn2018].
There is much less agreement regarding the less heavily worked and formally standardized LCTs typical of the earliest Acheulean [@beyene2013; @delatorre2018; @diez-martín2015; @lepre2011; @semaw2018].
Such forms continue to occur with variable frequency in later time periods [@mcnabb2015], and may be especially prevalent in eastern Asia [@li2021].
Although formal types have been recognized in the Early Acheulean and are commonly used to describe assemblages, many workers now see a continuum of morphological variation [@duke2021; @kuhn2020; @presnyakova2018] including the possibility that simple flake production remained an important [@shea2010] or even primary [@moore2016] purpose of Early Acheulean large core reduction.

Typologically, LCTs are differentiated from Mode 1 pebble cores on the basis of size (\>10cm) and shape (elongation and flattening) [e.g., @isaac1977].
This consistent production of large, flat, and elongated cores in the Achuelean has long been thought to reflect the pursuit of desired functional and ergonomic properties for hand-held cutting tools [@wynn2018].
Unplanned flaking can sometimes produce cores that fall into the LCT shape range [@moore2016] and this is one possible explanation of the relatively small "protobifaces" that occur in low frequencies in Oldowan assemblages [@isaac1997].
However, the Early Acheulean is clearly distinguished from the Oldowan by the production of larger artifacts necessitating the procurement and exploitation of larger raw material clasts.
Although studies of handaxe variation often focus on shape rather than size, this shift is an important aspect of artifact design with relevance to both production and function.

Production of larger tools was accomplished either through a novel process of detaching and working Large Flake Blanks (LFBs) from boulder cores or simply by using larger cobble and slab cores [@delatorre2018; @isaac1969; @semaw2018].
Both may involve similar flaking "strategies" (e.g., bifacial or multifacial exploitation) to those present in the Oldowan [@duke2021] but require more forceful percussion to detach larger flakes.
This increases the perceptual motor difficulty of the task [@stout2002] and in many cases may have been accomplished using different percussive techniques and supports [@semaw2009].
These new challenges would have increased raw material procurement [@shea2010] and learning costs [@pargeter2019] as well as the risk of serious injury [@gala2023] associated with tool production.
This strongly implies intentional pursuit of offsetting functional benefits related to size increase.
These likely included tool ergonomics and performance [@key2017] as well as flake generation, resharpening, and reuse potential [@shea2010].
Early Acheulean LCT production is thus widely seen as a part of shifting hominin behavioral ecological strategies including novel resources and mobility patterns [@linaresmatás2021; @rogers1994].

The degree of intentional design reflected in the shape of Early Acheulean LCTs is more difficult to determine.
For example, LFB production using a simple "least effort" bifacial/discoidal strategy will tend to generate predominantly elongated (side or end struck) flakes [@toth1982] whether or not this is an intentional design target.
Similarly, the difficulty of flaking relatively spherical cobbles [@toth1982] might bias initial clast selection and subsequent reduction toward flat and elongated shapes even in the absence of explicit design targets.
On the other hand, it has been argued that the shape of Early Acheulean LFBs was intentionally predetermined using core preparation techniques [@delatorre2018] and many researchers perceive efforts at intentional shaping in the organization of flake scars on Early Acheulean handaxes and picks [@beyene2013; @delatorre2018; @diez-martín2015; @duke2021; @lepre2011; @semaw2009].
To date, however, the identification of Early Acheulean shaping has generally relied on qualitative assessment by lithic analysts.
Such assessment may in fact be reliable, but is subject to concerns about potential selectivity, bias, and/or overinterpretation [@davidson2002; @moore2016].
Notably, a 3-dimensional morphometric (3DGM) study by Presnyakova, et al. [-@presnyakova2018] concluded that LCT shape variation in the Okote Member (\~1.4 mya) at Koobi Fora was largely driven by reduction intensity rather than different knapping strategies.
However, this study did not directly address the presence/absence of design targets constraining the observed range of variation.

In later Acheulean contexts, reduction intensity effects are commonly equated with resharpening and seen as an alternative to intentional form imposition [@mcpherron2000].
Across heavily-worked and relatively standardized LCT assemblages [e.g., @shipton2020], a *lack* of association between morphology and reduction intensity has be used as an argument-by-elimination for the presence of imposed morphological norms or "mental templates" [@garcía-medrano2019; @shipton2015; @shipton2023].
However, in the less heavily-worked and more heterogeneous assemblages typical of the early Acheulean [@kuhn2020], it is equally plausible that increasing reduction intensity would reflect degree of primary reduction rather than subsequent resharpening [@archer2010].
In this case, reduction intensity effects on morphology would have the opposite interpretation: more reduction should result in closer approximation of a desired form if such were present.
For example, Beyene, et al. [-@beyene2013] found that increasing flake scar counts were associated with increasing handaxe refinement through time at Konso, Ethiopia, which may reflect a more general trend in the African Acheulean [@shipton2018].

Interpretive approaches address this quandary by "reading" the organization of scars on individual pieces to infer intent, but an adequate method to objectively quantify these insights has yet to be developed.
Current measures of reduction intensity, such as the scar density index (SDI) [@clarkson2013], are designed to estimate total mass removed from a core and are reasonably effective [@lombao2023].
However, mass removal was not the objective of Paleolithic flaking.
Indeed, knapping efficiency is usually conceived as generating an outcome while *minimizing* required mass removal.
This is true whether the desired outcome is a useful flake, a rejuvenated edge, or a particular core morphology.
In simple flake production, mass removed is probably a good reflection of the completeness of exploitation ("exhaustion") of cores and may have implications for required skill [@toth1982; @pargeter2023] as well as raw material economy [@reeves2021; @shick1987].
However, in core shaping and resharpening, mass removal would typically represent an energetic and raw material cost to be minimized, and might even interfere with function [@keyHandaxeShapeVariation2019].
Without further information, relationships between artifact shape and reduction intensity are thus open to conflicting interpretations as evidence of intentional design or its absence.

Li, et al. [-@li2015] proposed a Flaked Area Index (FAI) as an alternative to SDI as a measure of reduction intensity, arguing that its validity is supported by an observed correlation (r=0.424) with SDI.
However, they also explain that "flaked area does not necessarily relate to the number of flake scars\...a small number of large scars can produce a large area of scar coverage, and conversely, a large number of small scars can produce a small area of scar coverage." (p. 6).
We suggest that what FAI actually captures is the spatial extent modifications to the surface of a core.
It is thus complementary to the measure of volume reduction provided by SDI and provides additional information to inform technological interpretations.
For example, a correlation between FAI and artifact form without any effect of SDI would suggest a focus on "least-effort" shape imposition whereas the opposite pattern would be consistent with relatively intense resharpening of spatially restricted areas on the core.
A lack of shape correlation with either measure would be expected for simple debitage with no morphological targets whereas a strong correlation with both would indicate a highly "designed" form achieved through extensive morphological and volumetric transformation.
In the current study we thus considered SDI and FAI together in order to evaluate evidence of intentional shaping in the early Acheulean of Gona.

## Measuring artifact form and modification

Three-dimensional scanning and geometric morphometric (3DGM) methods are increasingly common in the study of LCT form and reduction intensity [@shipton2015; @caruana2020; @presnyakova2018; @li2021; @archer2010; @li2015; @lycett2006].
These methods provide high-resolution, coordinate-based descriptions of artifact form including detailed information about whole object geometric relations that is not captured by conventional linear measures [@shott2010].
This includes measures of surface area used to compute both SDI and FAI measures [@clarkson2013; @li2015].
At the time of writing, however, 3D scans are available for only a small number of Gona artifacts, including 33 of the Oldowan and Acheulean flaked pieces used in this study (**Figure** \@ref(fig:fig1)).
Despite continuing improvements, 3DGM methods still impose additional costs in terms of data collection and processing time as well as required equipment, software, and training.
Importantly, 3DGM methods cannot be applied to pre-existing photographic and metric data sets [e.g., @marshall], including available data from Gona.
For this reason, and to better understand the relative costs and benefits of 3DGM more generally, we sought to test the degree to which conventional measurements can approximate 3DGM methods and produce reliable results by directly comparing our conventional measures with 3DGM analysis of the 33 available scans.

```{r fig1, echo=FALSE, out.width = '80%', fig.align='center', fig.cap="3D Artifact scans from Gona used in this study."}
knitr::include_graphics(here::here("figure/Fig1.png"))
```

For our study, we are specifically concerned with the accurate description of morphological variation and estimation of artifact surface and flaked areas.
With respect to morphology, we were encouraged by the fact that aspects of form variation relevant to our research question (i.e., core elongation, flattening, and pointedness) are relatively simple to describe using sparse data.
3DGM studies of LCTs collect vast amounts of shape data but may discard upward of 50% of the observed variation in order to focus on two or three interpretable principal components.
Across studies, these PCs consistently correspond to basic features like elongation, relative thickness, pointedness, and position of maximum thickness that also emerge from lower-resolution spatial data [@garcía-medrano2019; @archer2010; @lycett2006] and studies employing linear measures rather than spatial coordinates [@pargeter2019; @crompton1993].
There is less evidence that conventional methods can accurately estimate artifact surface and flaked areas.
Clarkson [-@clarkson2013] advocated the use of 3D surface area measures as more accurate than estimation from linear measures (e.g., surface area of a rectangular prism defined by artifact dimensions).
However, he also found that the error introduced by the linear approach was a highly systematic, isometric overestimation of surface area and that results correlated with direct 3D measures with an impressive r^2^ = 0.944 and no effect of variation in core shape.
Insofar as it is variation in the relationship between surface area and flaking intensity that is of interest, rather than absolute artifact size, such systematic overestimation may not be problematic.
Similar concerns apply to the estimation of flaked area.
Traditionally, such estimates have been done "by eye" as a percentage of the total artifact surface [e.g., @dibble2005].
Such estimations have been found to be reasonably accurate when compared to 3D methods, but with the potential for substantial error on individual artifacts [@lin2010].
The accuracy of by-eye estimation has yet to be systematically studied in Early Stone Age cores like those from Gona.

## The Early Acheulean at Gona

Early Acheulean sites in the Gona Project area are distributed over a wide area within the Dana Aoule North (DAN), Ounda Gona South (OGS), and Busidima North (BSN) drainages in the Busidima Formation [@quade2004] and range in age from approximately 1.7 to 1.2 mya [@semaw2018].
The specific sites included in the current analysis are DAN-5, OGS-5, OGS-12, and BSN-17, all estimated to ca.
1.7 -- 1.4 mya by stratigraphic position in the Gona sequence [@quade2008; @semaw2020].
The Busidima Formation accumulated through fluvial deposition by the ancestral Awash River (Type I context) and its smaller tributary channels (Type II context) [@quade2004; @quade2008].
Oldowan sites at Gona all occur in Type I sediments, indicating channel bank/margin (OGS-7) or proximal floodplain (EG-10, EG-12) contexts close to the large, hetero-lithic clasts available from point bars in the axial river channel [@quade2004; @stout2005].
Acheulean sites continue to occur in Type I contexts (BSN-17, DAN-5) but are also found in Type II sediments (OGS-5, OGS-12,) reflecting increased utilization of large perennial tributaries to the ancestral Awash River [@quade2008].
Clasts locally available in these tributaries were relatively small, implying that the large flakes and cobbles used to produce Acheulean artifacts were initially sourced from the axial river.
A similar pattern of habitat diversification and increasing lithic transport distances has been described at other sites and may be typical of the early Acheulean [@rogers1994; @linaresmatás2021; @hay1976].
As with other early (i.e. \>1.0 mya [@stoutStoneToolmakingEvolution2011; @presnyakova2018]) Acheulean assemblages, the Gona collections include numerous "crudely made" handaxes and picks on large flake blanks and cobbles, as well as large (\> 10cm) unmodified flakes, flaked pieces interpreted typo-technologically as Mode 1 cores (see Figure 1af), and smaller debitage [@semaw2018; @semaw2020].

# Materials and Methods

## Materials

### Archaeological Sample

Artifact collections analyzed here include *in situ* pieces excavated from intact stratigraphic contexts and surface pieces systematically collected from the sediments eroding from these layers.
Surface pieces are included because the current technological analysis does not require more precise spatial association, stratigraphic, and chronological control.
Our sample comprises the total collection of flaked pieces [@isaac1997] and large (\>10 cm) detached pieces from each site, regardless of typo-technological interpretation.

## Methods

### Artifact Classification

Artifacts were classified according to initial form (pebble/cobble, detached piece, or indeterminate), presence/absence of retouch, technological interpretation ("Mode 1" core vs. "Mode 2" LCT), and archaeological context (Oldowan vs. Early Acheulean sites).
LCTs were additionally classified as handaxes, knives, or picks following definitions from Kleindienst [-@kleindienst1962].
The validity of technological interpretations and typological classifications was assessed through cluster analysis based on artifact shape and reduction intensity variables.

### Artifact Measurement

Conventional linear measures capture the direction (e.g., length \> breadth) but not the location of geometric relations (e.g., position of maximum breadth).
We address this by collecting linear measures defined by homologous semi-landmarks.
All artifacts were oriented along their maximum dimension, which was measured and defined as "length." The next largest dimension orthogonal to length was used to define the plane of "breadth," with the dimension orthogonal to this plane defined as "thickness" Width (W~1~, W~2~, W~3~) and thickness (T~1~, T~2~, T~3~) measures were then collected at 25%, 50%, and 75% of length, oriented so that 25% Width \> 75% Width.
To partition variation in shape from variation in size, we divided all linear measures by the geometric mean [@lycett2006].
GM-transformed variables were then submitted to a Principal Components Analysis (covariance matrix) to identify the main dimensions of shape variation.

Our semi-landmark measurement system allowed us to improve on the prism-based surface area formula (2LW + 2LT + 2 WT) by using our 7 recorded dimensions to more tightly fit three prisms around the artifact: SA = W~1~T~1~ + 2(.33L \* W~1~) + 2(.33L \* T~1~) + 2(.33L \* W~2~) + 2(.33L \* T~2~) + 2(.33L \* W~3~) + 2(.33L \* T~3~) + W~3~T~3~.
Surface area calculated in this way correlates with mass^2/3^ at r^2^ = 0.947 in our sample.
Calculated surface area was then used to derive the Scar Density Index: SDI = number of flake scars \> 1cm per unit surface area [@clarkson2013; @shiptonFlakeScarDensity2015].
The Flaked Area Index (FAI: flaked area divided by total surface area)[@li2015] was estimated directly "by eye" as a percentage of the total artifact surface.

### 3D Methods

Artifact scans (N=33) were cleaned, smoothed, and re-meshed using MeshLab.
The 3D triangular mesh of each scan was computed using Kazhdan and Hoppe's [-@kazhdan2013] method of screened Poisson surface reconstruction in MeshLab.
3DGM analysis was conducted using the *AGMT3-D* program of Herzlinger and Grosman [-@herzlinger2018].
Artifacts were automatically oriented according to the axis of least asymmetry, then manually oriented in the interests of standardization with the length, width, and thickness dimensions defined by the longest axis, followed by the next two longest axes perpendicular to the first.
The wider end of the artifacts was positioned as the proximal end (base), and more protruding surfaces were oriented towards the user in the first orthogonal view.
Then, a grid of 200 homologous semi-landmarks were overlain on each artifact's surface.
Generalized Procrustes and Principal Component analyses were then undertaken to explore the shape variability of the sample.
The surface area of each artifact was calculated using the *Artifact3-D* program of Grosman et al. [-@grosman2022].
*Artifact3-D* was also used to automatically identify the flake scar boundaries and compute each scar's surface area, using the scar analysis functions of Richardson et al. [-@richardson2014].

### Analyses

Measurement Validation: To assess the adequacy of shape descriptions based on our linear measures, we directly compared these with shape as quantified by 3D methods on the 33 artifacts for which scans are available.
GM-transformed linear measures from these 33 artifacts were submitted to a variance-covariance matrix PCA.
PCs with an eigenvalue greater than the mean were retained for analysis and the results compared qualitatively (morphological interpretation of PCs) and quantitatively (correlation of artifact factor scores) to 3D results.
Accuracy of surface area and flaked area estimates was also assessed by correlation with 3D results.

# Results

## Measurement Validation

A PCA on GM transformed linear measures the 33 artifacts identified two PCs accounting for 75.6% of the variance.
Linear-PC1 (Eigenvalue = 0.189) explained 55.2% of the variance.
Factor loadings (**Table** \@ref(tab:tab1)) for Linear-PC1 reflect artifact elongation (i.e., an anti-correlation of length vs. distal width and thickness).
This A PCA on GM transformed linear measures the 33 artifacts identified two PCs accounting for 75.6% of the variance.
Linear-PC1 (Eigenvalue = 0.189) explained 55.2% of the variance.
Factor loadings (**Table** \@ref(tab:tab1)) for Linear-PC1 reflect artifact elongation (i.e., an anti-correlation of length vs. distal width and thickness).
This closely parallels the length vs. width and thickness tradeoff captured by 3DGM-PC1 (Figure 3a) and is reflected in a tight correlation of artifact scores produced by the two PCs (r = 0.903, p \< 0.001, Figure 3c).
Comparison with conventional shape ratios (Table 2) similarly indicates that both Linear-PC1 and 3DGM-PC1 largely capture variation in artifact elongation.
A second factor (Linear-PC2, eigenvalue = 0.07) explained an additional 20.4% of variance.
This factor was less strongly correlated with its 3DGM counterpart (3DGM-PC2; r = 0.344, p = 0.050) probably because Linear-PC2 describes anticorrelated variation in width and thickness (i.e., broad and flat vs. thick and pointed; Table 1) whereas 3DGM-PC2 more purely isolates convergence (Figure 3a).
The remainder of the shape variability explained by Linear-PC2 is captured by higher order 3DGM-PCs like 3DGM-PC3 to 5, which comprise the contribution of the left and right lateral margins to relative thickness.
Use of high-resolution, coordinate-based scan data thus generates PCs that identify more specific shape attributes, but the underlying morphological variability captured by the linear and 3D analyses remains similar.
Together, 3DGM-PC2 (r = 0.344, p = 0.050), 3DGM-PC3 (r = -0.416, p = 0.016), 3DGM-PC4 (r = 0.458, p = 0.007), and 3DGM-PC5 (r = -0.352, p = 0.044) correlate well with Linear-PC2 (Figure 3c), cumulatively capturing whether the items are broad and flat or thick and pointed.
A stepwise regression (r2=0.625, F(4,28)=11.697, p\<0.001, Probability-of-F-to-enter \<= 0.050; Probability-of-F-to-remove \>= 0.100) with Linear-PC2 as the dependent variable retained all four of these 3DGM-PCs as significant predictors.

```{r tab1, echo=FALSE, message=FALSE}
library(tidyverse)
df1 <- data.frame("Linear metrics (GM transformed)" = c("Length", "W1", "W2", "W3", "T1", "T2", "T3"), 
                 "Linear-PC1" = c("0.989", "0.303", "0.403", "-0.176", "-0.135", "-0.369", "-0.607"), 
                 "Linear-PC2" = c("-0.107", "0.350", "0.767", "0.790", "-0.679", "-0.623", "-0.282") ) 
kableExtra::kbl(df1, booktabs = T, caption = "Component loadings for linear metric PCs .") 
```

We thus concluded that linear measures are adequate to capture relevant shape variation and proceeded with a PCA on our full sample.
We identified two PCs accounting for 80.0% of the variance.
PC1 (Eigenvalue = 0.216) explained 56.4% of the variance.
Factor loadings (Table 2) for PC1 reflect artifact flatness (i.e., an anti-correlation of length and width vs. thickness) such that higher values indicate relatively thinner pieces.
PC2 (Eigenvalue = 0.090) explained 23.6% of the variance.
Factor loadings (Table 2) show that PC2 captures artifact convergence or "pointedness" (i.e., an anti-correlation of tip width with length and butt thickness) such that higher values indicate shorter, less pointed forms.

We also tested the validity of our two reduction measures, SDI and FAI.
In agreement with Clarkson (2013), we found that surface area estimated from caliper measures displayed a strong correlation with (r2=0.975, p \< 0.001) but linear over-estimation of ($\beta$ = 1.58) 3DGM surface area (Figure 4a).
This results in a systematic underestimation of SDI that scales with core size (Figure 4b).
However, a simple correction of the caliper estimate (dividing by the slope, 1.58) eliminates surface area over-estimation (mean difference = 256mm2 [\<1.7% of mean], p=0.040) and produces SDI values that agree with 3DGM (r2=0.975, p \< 0.001, $\beta$ = 0.98) (Figure 4c).
We thus proceeded to apply this correction to surface area estimates in the full sample.
Insofar as these relationships are driven by basic geometry, we expect these methods (including correction) to be generalizable to other ESA assemblages.

## Classification Validation

We first conducted a stepwise DFA on all flaked pieces (n=192) with inferred technological Mode (one vs. two) as the grouping variable and PCs 1 and 2, corrected SDI (cSDI), and FAI as the independent variables.
All variables were retained, yielding one canonical DF (eigenvalue=1.825, Wilks Lambda = 0.354, p\< 0.001) which correctly classified 93.8% of artifacts.
We thus accepted the validity of classification by Mode in our sample and employed this distinction in subsequent analyses.
There was no discernable difference in discriminant scores for Mode 1 cores from Oldowan (n=37) vs. Acheulean (n=39) contexts (, p = 0.746).
Mode 1 cores from Oldowan contexts (n=37) do include 10 (27%) small, retouched flakes.
Only one retouched flake from an Acheulean context was classified by the DFA as a Mode 1 core.
This piece, typologically classified as a "knife", is the smallest (93mm) retouched flake in the Acheulean sample.
When retouched flakes are excluded from the comparison, there are no significant differences in shape between Mode 1 cores from Oldowan and Achuelean contexts.
Interestingly, however, Acheulean Mode 1 cobble-cores are much larger (mean weight 480g vs. 186g, Cohen's d=1.137, p\<0.001) and less heavily reduced (mean cSDI 0.057 vs. 0.103, Cohen's d = -0.884, p \< 0.001) despite having similar FAI (mean 0.52 vs. 0.46, Cohen's d=0.271, p=0.266).

Next, we conducted a stepwise DFA on all flaked Mode 2 pieces (i.e. excluding unmodified large flakes, n = 115) with typology (handaxe, pick, knife) as the grouping variable and the same four independent variables.
Both shape PCs and FAI were retained, while cSDI was not entered.
This produced two DFs (DF1: Eigenvalue=1.536, 91.3% of variance; DF2: Eigenvalue = 0.146, 8.7% of variance; Wilks Lambda = 0.344; p\<0.001) which correctly classified 71.3% of artifacts.
Inspection of DF coefficients shows that DF1 captures an inverse correlation between flaked area and pointedness (Linear-PC2) whereas DF2 captures a positive relationship between flaked area and elongation (Linear-PC2).
A bivariate plot (Figure 5) illustrates the fact that DF1 captures a range of variation from pointed, heavily-worked picks, through handaxes, to knives, with substantial overlap between adjacent types (Table 2).
As an intermediate type, "handaxes" were correctly classified only 43.9% of the time.
In agreement with others [@duke2021; @kuhn2020; @presnyakova2018] we conclude that these typological labels artificially partition a continuum of variation and abandon them in subsequent analyses.

## Effects of Reduction on Shape

To assess the influence of flake removals on core form, we examined the association between our reduction measures (cSDI and FAI) and core shape (PC1, PC2).
In the complete sample of flaked pieces (n=192), we observed weak but significant effects of cSDI (r=-0.294, p \<0.001) and FAI (r=-0.294, p\<0.001) on PC1(flatness) and of FAI only on PC2(pointedness) (r=-0.436, p\<0.001).
However, it is clear that these overall effects conflate different trends in Mode 1 vs. Mode 2 cores, as well as in cores executed on flake vs. pebble/cobble bases.
Within categories, we observed no significant effects of cSDI whereas FAI had variable relationships with core form.

In Mode 1 cores, a weak negative effect of FAI on PC2(pointedness) (n = 76, $\beta$ =-0.006, r=0.240, p= 0.037) suggests that increasing extent of modification tends to slightly increase pointedness (Figure 6a).
No effects of FAI on PC1 (flatness) or of SDIc on either PC approach significance.
FAI and SDIc are moderately correlated ($\beta$=0.001, r=0.524, p\<0.001), an effect that reflects increases in the upper limit of scar density values as flaked area increases (i.e., flaked area constrains maximum cSDI: Figure 6b).

In Mode 2 cores (n=115), reduction measures have different effects depending on initial blank form.
In general, the contrast between flat, lightly-worked flake bases and rounder, more-heavily worked cobble bases tends to inflate relationships between FAI and core shape (Figure 7c,d).
Mode 2 cobble cores (n=37) display a negative effect of FAI on PC1(Flatness) ($\beta$=-0.011, r=0.367, p = 0.026), no effect of FAI on pointedness, and no association between FAI and cSDI (r=0.141, p = 0.406).
Flake cores (n=69) show negative effects of FAI on PC1(Flatness) ($\beta$=-0.007, r=0.284, p = 0.018) and, more strongly, PC2(pointedness) ($\beta$=-0.019, r=0.443, p \< 0.001).
Flake cores show a small but significant positive effect of FAI on cSDI that, in contrast to Mode 1 cores, represents actual covariance rather than simply a relaxation of constraint (i.e. both upper and lower limits of cSDI increase with FAI).

Comparison of Mode 2 flake cores with unmodified large flakes from Achuelean contexts (n=35) shows that ranges of shape variation substantially overlap but that flaking generally reduces both PC1(flatness) and PC2(pointedness).
In fact, regressions of PC1 and PC2 on FAI show y-intercepts that closely approximate unmodified flake mean values (0.852 vs. 0.774 and 0.693 vs. 0.594, respectively) and significantly negative slopes as FAI increases (Figure 9).

A small number of Mode 2 cores (n=9) were so heavily modified (mean FAI=89%, range 65-100) that the blank form could not be determined.
These heavily worked "indeterminate" cores tend to show similar PC1(flatness) and PC2(pointedness) values to cobble-base cores (i.e. thicker and more pointed than flake cores; Figure x).
They also appear to follow general trends of shape change with increasing FAI observed across base types (Figure 10).

# Discussion

Following the criteria we proposed in the Introduction, Large Flake base cores from Acheulean contexts show evidence of design.
Importantly, "design" in this context is understood as "design in four dimensions" sensu Kuhn [-@kuhn2020].
Increasing FAI is associated with progressive narrowing ( i.e., decreasing PC1) and convergence (decreasing PC2) whereas cSDI is not.
This indicates that the size and non-overlapping placement of flake removals drives shape change, rather than simply the removal of mass in general as might be expected from conventional models of resharpening (?).
The fact that cSDI nevertheless correlates with FAI is consistent with relatively light (number of scars per piece ranges from 2 to 17 with a mean of 6.94) and spatially distributed re-touching involving relatively little overlap between scars.
Such a pattern could be produced by intentional imposition of a design target or "mental template" or a more procedural/functional bias in favor of establishing cutting edges at one end of the long axis.
In the absence of effective thinning/volume management techniques, this will lead to rapid narrowing/thickening of the piece overall and especially at the "tip." This is consistent with differences between flaked vs. unmodified large flakes and furthers the point that blank production already can be conceived as a process of "design" -- many unmodified flakes may have been similar in shape and just as useful as many retouched tools.

This contrasts with reduction of Mode 1 cores.
Unexpectedly, FAI did have an effect to increase convergence.
However, this effect was substantially weaker than seen in LFB Acheulean cores and may simply reflect the fact that reduction is unlikely to be perfectly symmetrical and our orienting protocol guarantees that one end is always more pointed than the other unless they are precisely equal in breath.
In support of this "byproduct" interpretation, the relationship between FAI and cSDI is essentially what one would expect from knapping unconstrained by any systematic core-shaping strategy (greater flaked area can, but does not necessarily, accommodate a greater number of flake scars)

Our most complicated results come from Mode 2 cobble cores.
Evidence of intentional shaping is limited: FAI has a moderate effect on PC1(flatness) and no effect on pointedness.
Furthermore, the is no association between FAI and cSDI cobble cores display a greater range of cSDI values.
Instead, FAI is strongly correlated with scar size (mean scar length: n=33, r=.588, P\<0.001; Figure 11).
Together, this suggests that the removal of large flakes in particular increases FAI and tends to narrow the piece.
We propose that this pattern may actually indicate that at least some cobble cores were precisely that: cores for the generation of LFBs.
Given the difficulty of working relatively spherical cobbles, we expect that knappers would have chosen elongated and/or tabular cobbles.
Attempts to maximize potential flake dimensions would naturally lead to removals along the long axis rather than at either end.
This hypothesis was suggested by informal "experiential" knapping by the lead author with Gona cobbles, during which LFB production produced pick like forms as a byproduct.
In support of the plausibility of this hypothesis, the maximum scars length on Gona Mode 2 cobble cores substantially overlaps with the size of unmodified large flakes from Acheulean contexts.
It is also possible that cobble cores might begin as cores for LFB production and then be finished into core tools in their own right.
These possibilities have major implications for understanding the occurrence and apparent morphological stability [@beyene2013] of "picks" throughout the African Acheuelan but remains to be tested.

# References